---
output:
  html_document: default
  pdf_document: default
  word_document: default
---
2008 Presidential Campagin Finance for president Barack Obama by Alina Jurgensen
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.
#install.packages('ggplot2')
library(ggplot2)
#install.packages('knitr', repos='http://cran.rstudio.org')
#install.packages('ggplot2',repos = "http://cran.us.r-project.org")
suppressMessages(library(dplyr))
#install.packages('ggthemes', dependencies = TRUE) 
library(ggthemes)
#theme_set(theme_minimal(24)) 
library(plyr)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Load the Data
el <- read.delim('ElectionsHI.txt')
summary(el)

```

# Univariate Plots Section


My report explores the 2008 Presidential Campagin Finance dataset containing data about financial support to president Barack Obama, in his native state, Hawaii.




```{r echo=FALSE, message=FALSE, warning=FALSE}
 qplot(x = contbr_city, data = el, binwitdth =25) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "The cities with higher contributers", x = 'Contributers city')
```



The majority of countributors come from Honolulu, followed by Kailua and Hilo.



```{r echo=FALSE, message=FALSE, warning=FALSE}


county <- function(x) {
  if( is.na(x) ) { NA}
  else if (x ==  "HONOLULU" | x == "KAILUA KONA" | x == "KANEOHE" | x == "WAIANAE" | x == "MILIANI" | x == "KANEOHE" | x == "KAILUA" | x == "WAIALUA" | x == "KANEOHE" | x == " WAIPAHU" | x == "PEARL CITY" | x == "KAAAWA " | x == " HALEIWA " | x == "KAPOLEI" | x == "AIEA" | x == "??huimanu" | x == "Aiea"| x == "East Honolulu" | x == "Ewa Beach" | x == "Ewa Gentry" | x == "Ewa Villages" | x == "H??lawa" | x == "Hau??ula" | x == "He??eia" | x == "Iroquois Point" | x == "Ka??a??awa" | x == "Kahalu??u" | x == "Kahuku" | x == "Kalaeloa" | x == "K??ne??ohe" | x == "Kapolei" | x == "Kawela Bay" | x == "Ko Olina" | x == "L????ie" | x == "M????ili" | x == "M??kaha" | x == "M??kaha Valley" | x == "Makakilo" | x == "Maunawili" | x == "Mililani Mauka" | x == "Mililani Town" | x == "Mokul????ia" | x == "N??n??kuli" | x == "Ocean Pointe" | x == "Punalu??u" | x == "P??p??kea" | x == "Royal Kunia " | x == "Schofield Barracks" | x == "Wahiaw??" | x == "Wai??hole" | x == "Waialua" | x == "Wai??anae" | x == "Waik??ne" | x == "Waikele" | x == "Waimalu" | x == "Waim??nalo" | x == "Waim??nalo Beach" | x == "Waipahu" | x == "Waipi??o" | x == "Waipi??o Acres" | x == "West Loch Estate" | x == "Wheeler AFB" | x == "Whitmore Village")
    { "Honolulu"} 
  else if (x == "CAPTAIN COOK" | x == "KEAAU" | x == "HILO" | x == "HOLUALOA" | x == " KAMUELA" | x == "VOLCANO" | x == "WAIKOLOA" | x == "HONOKAA" | x == " HAWAII NATIONAL PA" | x == "Kailua" | x == "Kalaoa" | x == "Waimea" | x == "Holualoa" | x == "Waikoloa Village" | x == "Hawaiian Ocean View" | x == "Mountain View" | x == "	Kahaluu-Keauhou" | x == "Ainaloa" | x == "Honaunau-Napoopoo" | x == "Honalo" | x == "Honalo" | x == "Honokaa" | x == "Keaau" | x == "Kealakekua" | x == "Pepeekeo" | x == "Kapaau" | x == "Pahala" | x == "Papaikou" | x == "Kurtistown" | x == "Wainaku" | x == "Hawi" | x == "Pahoa" | x == "Naalehu" | x == "Puako" | x == "	Paauilo" | x == "Laupahoehoe" | x == "Honomu" | x == "Halaula" | x == "	Paukaa" | x == "Kukuihaele" | x == "Waiohinu")
    {"Hawaii"} 
  else if (x == " KALAHEO" | x == "LIHUE" | x == "KAPAA" | x == "KILAUEA" | x == "KOLOA" | x == "KAPAA" | x == "LAWAI " | x == "HANALEI" | x == "Hanamaulu" | x == "Kekaha" | x == "	Puhi" | x == "Hanapepe" | x == "Eleele" | x == "Lawai" | x == "Anahola" | x == "Wailua" | x == "Princeville" | x == "Princeville" | x == "Waimea" | x == "Waimea" | x == "Omao" | x == "Poipu" | x == "Kaumakani" | x == "Hanalei" | x == "Haena" | x == "	Kalihiwai" | x == "	Wainiha" | x == "Pakala Village" | x == "	Kahului" | x == "Kihei" ) 
  { "Kauai"} 
  else if (x == "KIHEI" | x == "HAIKU" | x == "PAIA" | x == "KULA" | x == "LAHAINA" | x == "KAHULUI" | x == "WAILUKU" | x == "HANA" | x == "MAKAWAO" | x == "LANAI CITY" | x == "Waihee-Waiehu" | x == "Haiku-Pauwela" | x == "Pukalani" | x == "Napili-Honokowai" | x == "Makawao" | x == "" | x == "Wailea" | x == "	Kaunakakai" | x == "Waikapu" | x == "	Kualapuu" | x == "Keokea" | x == "Olinda" | x == "Kaanapali" | x == "Haliimaile" | x == "	Mahinahina" | x == "Launiupoko" | x == "Ualapu'e" | x == "Maunaloa" | x == "Kapalua" | x == "Maalaea" | x == "Makena" | x == "Olowalu" | x == "Manele") {
    "Maui"
  }
  else{
    "Other"
  }
  
}


el$counties <- apply(el['contbr_city'], 1, county)
 qplot(x = el$counties, data = el) +
   labs(title = "Counts by county", x = 'County')

```


I create a new category 'County' and I see that most contributors come from Honolulu county.





```{r echo=FALSE, message=FALSE, warning=FALSE}

occupation <- function(x) {
  if( is.na(x) ) { NA}
  else if (x ==  "SELF EMPLOYED" | x == "BUSINESS OWNER" | x == "INNKEEPER-OWNER" | x == " SMALL BUSINESS OWNER" | x == "HOMEMAKER" | x == "ENTREPRENEUR" | x == "SMALL BUSINESS OWNER") 
    { "Self_employed"} 
  else if (x == "NOT EMPLOYED" | x == "HOMEMAKER") 
    {"Not_employed"} 
  else if (x == 'RETIRED') 
    {"Retired"}
  else if (x == "LAWYER" | x == "ATTORNEY" | x == "ATTORNEY AT LAW" | x == "GENERAL COUNSEL" | x == "PARALEGAL" | x == "PUBLIC RELATIONS" | x == "EXECUTIVE" | x == "OFFICE MANAGER" | x == "PLANNING FOR REAL ESTATE DEVELOPMENT" | x == "REAL ESTATE" | x == "ADMIN. MANAGER" | x == "IT CONSULTANT" | x == "PROJECT MANAGER" | x == " ANALYST" | x == "REAL ESTATE" | x == "FINANCIAL CONSULTANT" | x == "SERVICE COORDINATOR" | x == "PROGRAM COORDINATOR " | x == "PUBLIC RELATIONS AND MARKETING " | x == "CEO" | x == "ARCHITECT" | x == "REAL ESTATE BROKER" |  x == "PROJECT ASSISTANT" | x == " PROGRAMMER" | x == "ACCOUNT MANAGER " | x == "MEDIATOR " | x == "MARKETING" | x == "INSURANCE AGENT" | x == "EQUITY INVESTOR" | x == "FINANCIAL DIRECTOR") 
    { "Profesional and business services" } 
  else if (x == "SCHOOL ADMINISTRATOR" | x == "ASSISTANT PROFESSOR" | x == "TEACHER" | x == "YOUTH PROGRAM DIRECTOR" | x == "PROFESSOR " | x == " ART TEACHER" | x == "GRADUATE RESEARCH ASST." | x == "SUBSTITUTE TEACHER " | x == " MIDDLE SCHOOL SCIENCE TEACHER" | x == "TEACHER/MOM" | x == " PROFESSOR OF THEATRE" | x == "DIRECTOR OF TEACHING AND LEARNING" | x == "PROFESSOR EMERITUS" | x == " EDUCATION ADMINISTRATION" | x == "COLLEGE COUNSELOR")
    {"Educational services"}
  else if (x == "FACTORY")
    {"Manufacturing"}
  else if (x == "NUTRITIONIST" | x == "HEALTH CARE" | x == "DENTAL RESEARCHER" | x == "HEALTH ADMINISTRATOR" | x == "PHYSICIAN" | x == "CHILD CARE PROVIDER" | x == "RADIOLOGIST" | x == "NURSE PRACTITIONER" | x == "MASSAGE THERAPIST" | x == "SPEECH PATHOLOGIST" | x == "DENTIST") 
    {"Health care and social assistance"}
  else if (x == "WRITER/EDITOR" | x == "WRITER-EDITOR" | x == "DOCUMENTARY FILMMAKER" | x == "PUBLISHER" | x == " PHOTOGRAPHER" | x == "THEATRE PROMOTIONS" | x == "PHOTO JOURNALIST" | x == "ARTIST " | x == "TV PRODUCER" | x == "MOVIE MAKER") 
    {"Media"}
  else if (x == "CHEF" | x == "COFFEE ROASTER" | x == "BARTENDER/WAITRESS" | x == "INNKEEPER-OWNER " | x == "INKEEPER" | x == "TOURISM INDUSTRY" | x == "VACATION RENTAL") {"Leisure and hospitality"}
  else if (x == "FORENSIC SPECIALIST" | x == "COMMANDER, ECHO COMPANY, 2-25 AVIATION" | x == "U.S. MARINES " | x == "SOCIAL WORKER " | x == "SOLDIER") {
    "State and local government"
  }
  else{
    "Other"
  }
  
}


el$occup <- apply(el['contbr_occupation'], 1, occupation)
 qplot(x = el$occup, data = el) +
   labs(title = "Counts by contributior's occupation", x = 'Contributior occupation') +
   theme(axis.text.x = element_text(angle = 90, hjust = 1))

```



```{r echo=FALSE, message=FALSE, warning=FALSE}
library(gridExtra) 

 q1<- qplot(x = contb_receipt_amt, data = el) +
   labs(title='Contribution amount')
   #scale_x_continuous(limits = c(500, 1500), breaks = seq(0, 1500, 100)) +
   


q2<-  qplot(x = 100, y = contb_receipt_amt, data = el) +
   geom_boxplot() +
    coord_cartesian(ylim= c(0,150))

grid.arrange(q1, q2)
```

Most of the contributions are around 100, with 50% of the contributions greater than 75.







```{r echo=FALSE, message=FALSE, warning=FALSE}
  qplot(x = election_tp, data = el) +
  labs(title='Election type')

```


The primary election had received more contributions than the general.



# Univariate Analysis

### What is the structure of your dataset?

My dataset has 15243 observations and 18 variables. The most relevant variables for my report are:
- contbr_nm: the contributer's name
- contbr_city: the contributer's city
- contb_receipt_amt: the amount 
- contbr_occupation
- contbr_employer
- election_tp: election type



### What is/are the main feature(s) of interest in your dataset?
 
I would like to see the received financial support by looking at contributor's location, so I choose two counties to explore: Honolulu and Hawaii. The variables used are contbr_city  contb_receipt_amt and contbr_occupation.

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?
I will use election_tp to compare the data between the two election types.



### Did you create any new variables from existing variables in the dataset?

I created new categories for county, employer and occupation.

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?


# Bivariate Plots Section


```{r echo=FALSE, message=FALSE, warning=FALSE}
 qplot(x = contb_receipt_amt, data = el) +
  facet_wrap(~el$counties) 
  labs(title='Contribution amount by county') 

```

Due to their contrast, I chose to focus on 2 different counties: Honolulu and Hawaii. 




```{r echo=FALSE, message=FALSE, warning=FALSE}



Hawaii <- subset(el, contbr_city == "CAPTAIN COOK" | contbr_city == "KEAAU" | contbr_city == "HILO" | contbr_city == "HOLUALOA" | contbr_city == " KAMUELA" | contbr_city == "VOLCANO" | contbr_city == "WAIKOLOA" | contbr_city == "HONOKAA" | contbr_city == " HAWAII NATIONAL PA"  | contbr_city == "Kailua" | contbr_city == "Kalaoa" | contbr_city == "Waimea" | contbr_city == "Holualoa" | contbr_city == "Waikoloa Village" | contbr_city == "Hawaiian Ocean View" | contbr_city == "Mountain View" | contbr_city == "	Kahaluu-Keauhou" | contbr_city == "Ainaloa" | contbr_city == "Honaunau-Napoopoo" | contbr_city == "Honalo" | contbr_city == "Honalo" | contbr_city == "Honokaa" | contbr_city == "Keaau" | contbr_city == "Kealakekua" | contbr_city == "Pepeekeo" | contbr_city == "Kapaau" | contbr_city == "Pahala" | contbr_city == "Papaikou" | contbr_city == "Kurtistown" | contbr_city == "Wainaku" | contbr_city == "Hawi" | contbr_city == "Pahoa" | contbr_city == "Naalehu" | contbr_city == "Puako" | contbr_city == "	Paauilo" | contbr_city == "Laupahoehoe" | contbr_city == "Honomu" | contbr_city == "Halaula" | contbr_city == "	Paukaa" | contbr_city == "Kukuihaele" | contbr_city == "Waiohinu")

ggplot(aes(x = contb_receipt_amt, y = contbr_city), data= Hawaii) +
  geom_jitter() +
  labs(title='Contributors amount in Hawaii county')


```



```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = contb_receipt_amt, y = contbr_city), data= subset(el,  contbr_city == "CAPTAIN COOK" | contbr_city == "KEAAU" | contbr_city == "HILO" | contbr_city == "HOLUALOA" | contbr_city == " KAMUELA" | contbr_city == "VOLCANO" | contbr_city == "WAIKOLOA" | contbr_city == "HONOKAA" | contbr_city == " HAWAII NATIONAL PA")) +
  geom_jitter()+
  scale_x_log10() +
  labs(title= 'Contribution amount for Hawaii count')

```


```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(aes(x = contb_receipt_amt, y = contbr_city), data= Hawaii) +
geom_jitter(aes(color = election_tp), stat = 'summary') +
  scale_x_log10() +
  
  labs (title= 'Contributors amount in Hawaii county by election type')

```


```{r echo=FALSE, message=FALSE, warning=FALSE}

Honolulu <- subset(el, contbr_city ==  "HONOLULU" | contbr_city == "KAILUA KONA" | contbr_city == "KANEOHE" | contbr_city == "WAIANAE" | contbr_city == "MILIANI" | contbr_city == "KANEOHE" | contbr_city == "KAILUA" | contbr_city == "WAIALUA" | contbr_city == "KANEOHE" | contbr_city == " WAIPAHU" | contbr_city == "PEARL CITY" | contbr_city == "KAAAWA " | contbr_city == " HALEIWA " | contbr_city == "KAPOLEI" | contbr_city == "AIEA" | contbr_city == "??huimanu" | contbr_city == "Aiea"| contbr_city == "East Honolulu" | contbr_city == "Ewa Beach" | contbr_city == "Ewa Gentry" | contbr_city == "Ewa Villages" | contbr_city == "H??lawa" | contbr_city == "Hau??ula" | contbr_city == "He??eia" | contbr_city == "Iroquois Point" | contbr_city == "Ka??a??awa" | contbr_city == "Kahalu??u" | contbr_city == "Kahuku" | contbr_city == "Kalaeloa" | contbr_city == "K??ne??ohe" | contbr_city == "Kapolei" | contbr_city == "Kawela Bay" | contbr_city == "Ko Olina" | contbr_city == "L????ie" | contbr_city == "M????ili" | contbr_city == "M??kaha" | contbr_city == "M??kaha Valley" | contbr_city == "Makakilo" | contbr_city == "Maunawili" | contbr_city == "Mililani Mauka" | contbr_city == "Mililani Town" | contbr_city == "Mokul????ia" | contbr_city == "N??n??kuli" | contbr_city == "Ocean Pointe" | contbr_city == "Punalu??u" | contbr_city == "P??p??kea" | contbr_city == "Royal Kunia " | contbr_city == "Schofield Barracks" | contbr_city == "Wahiaw??" | contbr_city == "Wai??hole" | contbr_city == "Waialua" | contbr_city == "Wai??anae" | contbr_city == "Waik??ne" | contbr_city == "Waikele" | contbr_city == "Waimalu" | contbr_city == "Waim??nalo" | contbr_city == "Waim??nalo Beach" | contbr_city == "Waipahu" | contbr_city == "Waipi??o" | contbr_city == "Waipi??o Acres" | contbr_city == "West Loch Estate" | contbr_city == "Wheeler AFB" | contbr_city == "Whitmore Village")
   

ggplot(aes(x = contb_receipt_amt, y = contbr_city), data= Honolulu) +
  geom_jitter() +
  labs(title='Contributors amount in Honolulu county')


```





```{r echo=FALSE, message=FALSE, warning=FALSE}


ggplot(aes(x = contb_receipt_amt, y = contbr_city), data= Honolulu) +
  geom_jitter() +
    scale_x_log10() +
  labs(title= 'Contributors amount for Honolulu county, log10')



```



```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(aes(x = contb_receipt_amt, y = contbr_city), data= Honolulu) +
geom_jitter(aes(color = election_tp), stat = 'summary') +
  scale_x_log10() +
    labs(title= 'Contributors amount for Honolulu county, by election type')

```



```{r echo=FALSE, message=FALSE, warning=FALSE}


qplot(x = contbr_occupation, data= Hawaii) +
   theme(axis.text.x = element_text(angle = 90, hjust = 1)) +

  
  labs(title= 'Contributors occupation for Hawaii county')

summary(Hawaii$contbr_occupation)


```




```{r echo=FALSE, message=FALSE, warning=FALSE}


summary(Honolulu$contbr_occupation)

```


Running a summary for the occupation type, I see that for Hawaii county the top 5 most present are Retired, Drafting service, Professor, Not Employed and Teacher. 
While for Honolulu are: Retired, Attorney, Professor, Physician and Not Employed.

Further I will explore the data looking at these professions in particular.






```{r echo=FALSE, message=FALSE, warning=FALSE}

qplot(x = contbr_occupation, data = subset(Hawaii, contbr_occupation == 'RETIRED' | contbr_occupation == 'PROFESSOR' |contbr_occupation == 'TEACHER' | contbr_occupation == 'DRAFTING SERVICE' | contbr_occupation == 'NOT EMPLOYED'))+
      
  facet_wrap(~ contbr_city)+
  labs(title='Counts of contributions from top 5 occupation types, for cities in Hawaii county') + theme(axis.text.x = element_text(angle = 90, hjust = 1))

```




```{r echo=FALSE, message=FALSE, warning=FALSE}

qplot(x = contbr_occupation, data = subset(Honolulu, contbr_occupation == 'RETIRED' | contbr_occupation == 'PROFESSOR' |contbr_occupation == 'ATTORNEY' | contbr_occupation == 'PHYSICIAN' | contbr_occupation == 'NOT EMPLOYED'))+
      
  facet_wrap(~ contbr_city)+
  labs(title='Counts of contributions from top 5 occupation types, for cities in Honolulu county') +

 theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


For both counties and in most cities, Retired group are the highest contributors. In Hawaii county this is followed by Drafting service and in Honolulu by Attorney.




```{r echo=FALSE, message=FALSE, warning=FALSE}

qplot(x = contb_receipt_amt, y = ..count../sum(..count..), data = Hawaii,
      xlab = "Amount",
      ylab = "Proportion of the contribution",
      
      geom = 'freqpoly', color = election_tp) +
  scale_x_continuous(lim = c(0, 1500), breaks = seq(0, 1500, 250)) +



  
  labs(title= ' Proportion of contributors amount by election type, for Hawaii county')



```

```{r echo=FALSE, message=FALSE, warning=FALSE}

qplot(x = contb_receipt_amt, y = ..count../sum(..count..), data = Honolulu,
       xlab = "Amount",
      ylab = "Proportion of the contribution",
     
      geom = 'freqpoly', color = election_tp) +
  scale_x_continuous(lim = c(0, 1500), breaks = seq(0, 1500, 250)) +


  
    labs(title= ' Proportion of contributors amount by election type, for Honolulu county')



```


The primary election has a higher percentage of contributions. For Hawaii county there is a higher percentage for the general election for 250 amount.



# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?
The countribution amount varies in relation to location: cities with higher population (Hilo, Honolulu) have more contributers and I was expected that. 


### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

There is a strong relation between contb_receipt_am and election_tp: the primary election has a higher percentage of small amounts.


### What was the strongest relationship you found?

Although I knew that Hawaii is the favorite place of retirment for many, I was surprised to see this group so strong in larger cities(Hilo, Honolulu)  and such a strong relation between contbr_occupation and contb_receipt_amt.



# Multivariate Plots Section




```{r echo=FALSE, message=FALSE, warning=FALSE}

qplot(x = contb_receipt_amt, y = contbr_occupation, data = subset(Honolulu, contbr_occupation == 'RETIRED' | contbr_occupation == 'ATTORNEY' | contbr_occupation == 'PROFESSOR' | contbr_occupation == 'NOT EMPLOYED'| contbr_occupation == 'PHYSICIAN'),
     
      color = election_tp) +
    
  scale_x_continuous(lim = c(0, 1000), breaks = seq(0, 1000, 100)) +
  geom_jitter() +
    labs(title= 'Contributions by top 5 occupation types for Honolulu county')




```



```{r echo=FALSE, message=FALSE, warning=FALSE}

qplot(x = contb_receipt_amt, y = contbr_occupation, data = subset(Hawaii, contbr_occupation == 'RETIRED' | contbr_occupation == 'PROFESSOR' | contbr_occupation == 'DRAFTING SERVICE' | contbr_occupation == 'NOT EMPLOYED'| contbr_occupation == 'TEACHER' ),
      xlab = ' amt',
      ylab = 'Proportion ',
      color = election_tp) +
     # geom = 'freqpoly', color = contbr_city) +
  scale_x_continuous(lim = c(0, 1000), breaks = seq(0, 1000, 100)) +
  geom_jitter() +
    labs(title= 'Contributions by top 5 occupation types for Hawaii county')




```





```{r echo=FALSE, message=FALSE, warning=FALSE}

qplot(x = contb_receipt_amt, y = contbr_occupation, data = subset(Hawaii, contbr_occupation == 'RETIRED' | contbr_occupation == 'PROFESSOR' | contbr_occupation == 'DRAFTING SERVICE' | contbr_occupation == 'NOT EMPLOYED'| contbr_occupation == 'TEACHER'  ),
      
     color = contbr_city) +
  scale_x_continuous(lim = c(0, 1000)) +


  
  labs(title= 'Contributions by top 5 occupation types for Hawaii county')



```






```{r echo=FALSE, message=FALSE, warning=FALSE}

qplot(x = contb_receipt_amt, y = contbr_occupation, data = subset(Honolulu, contbr_occupation == 'RETIRED' | contbr_occupation == 'ATTORNEY' | contbr_occupation == 'PROFESSOR' | contbr_occupation == 'NOT EMPLOYED'| contbr_occupation == 'PHYSICIAN'  ),
      
     color = contbr_city) +
  scale_x_continuous(lim = c(0, 1000), breaks = seq(0, 1000, 100)) +


  
  labs(title= 'Contributions by top 5 occupation types for Honolulu county')



```


### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?



### Were there any interesting or surprising interactions between features?
 

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

------

# Final Plots and Summary








### Plot One



```{r echo=FALSE, message=FALSE, warning=FALSE}
library(gridExtra) 

 q1<- qplot(x = 100, y = contb_receipt_amt, data = Hawaii) +
   geom_boxplot() +
  coord_cartesian(ylim=c(0,150)) +
   labs(title="Hawaii county")
   


q2<-  qplot(x = 100, y = contb_receipt_amt, data = Honolulu) +
   geom_boxplot() +
  coord_cartesian(ylim=c(0,250)) +
  labs(title="Honolulu county")

grid.arrange(q1, q2)
```


### Description One

Most of the contributions are amount of  75 and 100. for Honolulu county 25% of contributions are over 200 while for Hawaii are 100.





### Plot two



```{r echo=FALSE, message=FALSE, warning=FALSE}


q1<- qplot(x = contbr_occupation, data = subset(Hawaii, contbr_occupation == 'RETIRED' | contbr_occupation == 'PROFESSOR' | contbr_occupation == 'DRAFTING SERVICE' | contbr_occupation == 'NOT EMPLOYED'| contbr_occupation == 'TEACHER' && count(el$contbr_nm) > 20)) +
   labs (title= "Most loyal contributors, over 20 donations, Hawaii vs Honolulu")

q2<- qplot(x = contbr_occupation, data = subset(Honolulu, contbr_occupation == 'RETIRED' | contbr_occupation == 'PROFESSOR' | contbr_occupation == 'ATTORNEY' | contbr_occupation == 'NOT EMPLOYED' | contbr_occupation == 'PHYSICIAN'  && count(el$contbr_nm) > 20)) +
  labs("Honolulu county")

grid.arrange(q1,q2) 


```



### Description two

Looking at the most loyal financial supporters, by subsetting the data for individuals making over 20 contributions, I found that the two counties have in common three occupations: retired as the largest, followed by not-employed and professor.







### Plot Three



```{r echo=FALSE, message=FALSE, warning=FALSE}

q1 <- qplot(x = contb_receipt_amt, y = contbr_occupation, data = subset(Hawaii, contbr_occupation == 'RETIRED' | contbr_occupation == 'PROFESSOR' | contbr_occupation == 'DRAFTING SERVICE' | contbr_occupation == 'NOT EMPLOYED'| contbr_occupation == 'TEACHER'),
     
     color = contbr_city) +
  scale_x_continuous(lim = c(1500, 3000), breaks = seq(1500, 3000, 500)) +


  
  labs(title= 'Most generous contributors in Hawaii county') +
  facet_wrap(~election_tp)




q2 <- qplot(x = contb_receipt_amt, y = contbr_occupation, data = subset(Honolulu, contbr_occupation == 'RETIRED' | contbr_occupation == 'PROFESSOR' | contbr_occupation == 'ATTORNEY' | contbr_occupation == 'NOT EMPLOYED' | contbr_occupation == 'PHYSICIAN'),
     
     color = contbr_city) +
  scale_x_continuous(lim = c(1500, 3000), breaks = seq(1500, 3000, 500)) +


  
  labs(title= 'Most generous contributors in Honolulu county') +
  facet_wrap(~election_tp)

grid.arrange(q1,q2)


```


### Description Three

Looking at the most generous contributions, these come from Honolulu county. 
For Hawaii county, they are part of the retired group and come from Captain Cook city - which is not one of the largest cities. 





### Reflection
